A Technique for Inter-Catalogue Recommendations

### 2014-06-11 paper2/paper.tex
Rework ContentBasedRecommender subsection.

\section*{Experiments}
Our goal is to develop methods for connecting initially disconnected
recommendation graphs. Given two online retailers, each with a website
displaying their items along with several other {\em recommended} items from
their catalogue, we are presented with two disconnected recommendation graphs.
We would like to associate ``good'' recommendations for items in one catalogue
with items in the other catalogue. In this way, we are effectively introducing
edges that connect the two recommendation graphs. Formulating this
inter-retailer recommendations problem in terms of graphs offers insight in how
to evaluate our recommendation choices as well as how to choose ``good''
recommendations.  Leveraging the graph structure for evaluation is discussed in
this section while exploiting the graph for the purpose of making better
recommendations is addressed in a subsequent section.

@@

\section*{Constructing Recommendations}
Recommending products between different catalogues is a difficult task.
Typically, recommendations are constructed by leveraging user behavior
with a collaborative filtering technique. At the most basic level, If a large
number of users view or purchase both Product A and Product B, the two products
are recommended for each other. Unfortunately, for this problem there are no
user data between the two unconnected catalogues.

\subsection*{Content Based Similarity}
A simple approach that does not rely on user data is to recommend products that
have similar content. Each product page includes a short description which we 
can use to construct a textual content-based similarity function. To do this, 
we represent each product as a vector of tf-idf values for each word in its 
description. By using tf-idf, we reduce the relevance of frequent words and 
boost that of rare ones. This emphasizes the words that better represent this 
product versus all the other ones in the catalogue. The vector representations 
are the length of the vocabulary. These are sparse vectors as most words do 
not appear in any given product description. With this tf-idf represenation,
we can define the $ContentSimiarity$ functoin as follows:

\begin{align}
ContentSimilarity(P_1, P_2) &= v_1 \cdot v_2
\end{align}
where $P_1$ and $P_2$ are the products being compared and $v_1$ and $v_2$ are
their representations as tf-idf vectors.

This similarity function yields the following recommendation algorithm which we
call the $ContentBased Recommender$: For each product, calculate its
$ContentSimilarity$ score with each of the products in the other catalogue.
Predict recommendations between this product and the top 2 most similar products 
in the other catalogue.

While this is a quadric-time operation, it can be executed efficiently using 
matrix multiplication. Furthermore, the division into different categories 
reduces the impact of the quadratic effect. In our case, the entire Macy's 
catalogue consists of over $60,000$ products, making a quadratic time algorithm 
slow. However, the average category has only one to two thousand products, which
is a significantly more manageable size for quadratic operations.

>>>>

\section*{Experiment}
Our goal is to develop methods for connecting initially disconnected
recommendation graphs. Given two online retailers, each with a website
displaying their items along with several other {\em recommended} items from
their catalogue, we are presented with two disconnected recommendation graphs.
We would like to associate ``good'' recommendations for items in one catalogue
with items in the other catalogue. In this way, we are effectively introducing
edges that connect the two recommendation graphs. Formulating this
inter-retailer recommendations problem in terms of graphs offers insight in how
to evaluate our recommendation choices as well as how to choose ``good''
recommendations.  Leveraging the graph structure for evaluation is discussed in
this section while exploiting the graph for the purpose of making better
recommendations is addressed in a subsequent section.

@@

\section*{Algorithms}

\subsection*{\em ContentBasedRecommender}
The most straightforward approach to inter-retailer recommendations is a
content-based approach. Retail products typically have a description that can be
used to construct a textual content-based similarity function. One can then
consider all pairwise product combinations across retailers, but within a common
category, and choose recommendations between items with sufficiently high
similarity.

Toward this end, we borrow the concept of {\em Term Frequency--Inverse Document
Frequency} (tfidf) from the information retrieval community. In using tf-idf, we
are considering each item description to be a ``document''.  Thus we define the
tfidf of a term $t$ within an item description $d$ that's part of a combined
category catalogue $C$ (i.e. the catalogue consisting of all items in both
retailers' catalogues for a particular common category) as
\begin{align}
\mathrm{tfidf}(t,d,C) = f(t,d) \times
\mathrm{log}\frac{\left|C\right|}{\left|\left\{d \in C : t \in
d\right\}\right|},
\end{align}
where $f(t,d)$ is the number of occurances of term $t$ in item description $d$,
and $\left|\left\{d \in C : t \in d\right\}\right|$ is the number of items in
$C$ with descriptions containing at least one instance of term $t$.  We can now
represent each item as a sparse vector of tfidf values, which has a nonzero
element for each term in its description.

A key feature of tfidf that is useful here is how terms that appear in many item
description are discounted via the idf factor. Conversely, rare terms will have
larger tfidf values. We now define the $ContentSimilarity$ between items $A$ and
$B$ as
\begin{align}
ContentSimilarity(A,B) = v_A \cdot v_B
\end{align}
where $v_A$ and $v_B$ are the tfidf vector representations of items $A$ and $B$,
respectively.

Using this content-based similarity function, we construct the following
inter-catalogue recommendation algorithm, which we call the $ContentBased
Recommender$. For each product in each catalogue, calculate its
$ContentSimilarity$ with each and every product in the other catalogue. Predict
recommendations between this product and the top two most similar products in
the other catalogue. The reason for choosing the top two most similar products
originates from our knowledge that most items listed on www.macy.com are
associated with four recommended items. Thus, by choosing two edges per item per
partition, we will predict approximately the same number of edges that were lost
during the partition. This relatively straightforward algorithm serves as a
baseline against which to compare the performance of other more sophisticated
algorithms.

The $ContentBasedRecommender$ deserves a few more comments. For starters, the
use of an inner product, which is bounded only by the length of the tfidf
vectors, is not fundamental to this algorithm. A bounded similarity function
such as cosine similarity could have been used instead without qualitatively
changing the results discussed later in this report. This is because the lengths
of the product descriptions have little variation, and therefore, the vector
norms of the tfidf vectors have little variation. Finally, we mention that while
this algorithm is quadratic in the number items, it can be executed relatively
efficiently using matrix multiplication, which trades time for space (i.e.
memory). Furthermore, confining our experiments to individual categories reduces
the impact of the quadratic time complexity. In our case, Macy's entire
catalogue consists of over $60,000$ products, making a quadratic time algorithm
challenging. However, the average category has about two thousand products,
which is a significantly more manageable.
